---
title: Contingency analysis
subtitle: Comparing proportions among groups
bibliography: ../references.bib
---

<!-- COMMENT NOT SHOW IN ANY OUTPUT: Code chunk below sets overall defaults for .qmd file; these inlcude showing output by default and looking for files relative to .Rpoj file, not .qmd file, which makes putting filesin different folders easier  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

To introduce students to tests for two-samples and associated hypotheses
via contingency analysis. By the end of the lesson, you should be able
to

-   define and carry out contingency analysis

## Background reading

Course notes links for background reading also contain code used to
produce R output used in slides.

-   <https://jsgosnell.github.io/cuny_biostats_book/content/chapters/Compare_proportions_among_populations.html>

## [Lecture slides (click to open in Google slides!)](https://docs.google.com/presentation/d/1xsYPCNVGhV0RA33x6XtF0i1zhq9uWw2Qv08Iy3dptNM/edit?usp=sharing){target="_blank"}

```{=html}
<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQykVxAo_2bCH5IK1EH4Ix4Zq5LtCo1S4wtq-RNgLivFCInG_2PerwQXG5zmGEEI_T2MPol6m4xgjcL/embed?start=false&loop=false&delayms=60000" frameborder="0" width="640" height="389" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
```
## Connected swirl lesson

Swirl is an R package that provides guided lessons to help you learn and
review material. These lessons should serve as a bridge between all the
code provided in the slides and background reading and the key functions
and concepts from each lesson. A full course lesson (all lessons
combined) can also be downloaded using the following instructions.

**THIS IS ONE OF THE FEW TIMES I RECOMMEND WORKING DIRECTLY IN THE
CONSOLE! THERE IS NO NEED TO DEVELOP A SCRIPT FOR THESE INTERACTIVE
SESSIONS, THOUGH YOU CAN!**

-   install the "swirl" package

-   run the following code once on the computer to install a new course

    ```{r, eval=FALSE}
    library(swirl)
    install_course_github("jsgosnell", "JSG_swirl_lessons")
    ```

-   start swirl!

    ```{r, eval=F}
    swirl()
    ```

    -   swirl()

-   then follow the on-screen prompts to select the JSG_swirl_lessons
    course and the lessons you want

    -   **Here we will focus on the lessons on Contingency Analysis**

-   TIP: If you are seeing duplicate courses (or odd versions of each),
    you can clear all courses and then re-download the courses by

    -   exit swirl using escape key or bye() function

    -   uninstall_all_courses()

    -   then reinstall courses as needed

    -   when you restart swirl with swirl(), you may need to select

        -   2\. No. Let me start something new

## [Connected assignment(click here)](https://jsgosnell.github.io/cuny_biostats_book/content/practice_problems/5_Contingency_analysis.html){target="_blank"}

## Extra material

-   @davies1998

## Data referenced in class

-   @klem1989
