---
title: Compare proportions among groups
subtitle: Contingency analysis
bibliography: ../references.bib
---

<!-- COMMENT NOT SHOW IN ANY OUTPUT: Code chunk below sets overall defaults for .qmd file; these inlcude showing output by default and looking for files relative to .Rpoj file, not .qmd file, which makes putting filesin different folders easier  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

To introduce students to tests the idea of comparing groups and associated hypotheses
via contingency analysis. By the end of the lesson, you should be able
to

-   define and carry out contingency analysis
-   produce null and alternative hypotheses for comparing 2+ populations
-   define the issue of family-wise error rate and possible solutions

## Background reading

Course notes links for background reading also contain code used to
produce R output used in slides.

-   [https://jsgosnell.github.io/cuny_biostats_book/content/chapters/Compare_proportions_among_populations.html](https://jsgosnell.github.io/cuny_biostats_book/content/chapters/Compare_proportions_among_populations.html){target="_blank"}

## [Lecture slides (click to open in Google slides!)](https://docs.google.com/presentation/d/1xsYPCNVGhV0RA33x6XtF0i1zhq9uWw2Qv08Iy3dptNM/edit?usp=sharing){target="_blank"}

```{=html}
<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQykVxAo_2bCH5IK1EH4Ix4Zq5LtCo1S4wtq-RNgLivFCInG_2PerwQXG5zmGEEI_T2MPol6m4xgjcL/embed?start=false&loop=false&delayms=60000" frameborder="0" width="640" height="389" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
```
## Connected swirl lesson

Swirl is an R package that provides guided lessons to help you learn and
review material. These lessons should serve as a bridge between all the
code provided in the slides and background reading and the key functions
and concepts from each lesson. A full course lesson (all lessons
combined) can also be downloaded using the following instructions.

**THIS IS ONE OF THE FEW TIMES I RECOMMEND WORKING DIRECTLY IN THE
CONSOLE! THERE IS NO NEED TO DEVELOP A SCRIPT FOR THESE INTERACTIVE
SESSIONS, THOUGH YOU CAN!**

-   install the "swirl" package

-   run the following code once on the computer to install a new course

    ```{r, eval=FALSE}
    library(swirl)
    install_course_github("jsgosnell", "JSG_swirl_lessons")
    ```

-   start swirl!

    ```{r, eval=F}
    swirl()
    ```

    -   swirl()

-   then follow the on-screen prompts to select the JSG_swirl_lessons
    course and the lessons you want

    -   Here we will focus on the **Compare proportions among groups**
        lesson

-   TIP: If you are seeing duplicate courses (or odd versions of each),
    you can clear all courses and then re-download the courses by

    -   exiting swirl using escape key or bye() function

        ```{r, eval=F}
        bye()
        ```

    -   uninstalling and reinstalling courses

        ```{r, eval=F}
        uninstall_all_courses()
        install_course_github("jsgosnell", "JSG_swirl_lessons")
        ```

    -   when you restart swirl with swirl(), you may need to select

        -   No. Let me start something new

## [Connected assignment(click here)](https://jsgosnell.github.io/cuny_biostats_book/content/practice_problems/5_Contingency_analysis.html){target="_blank"}

Using these skills and applying concepts correctly to interpret data
sets may seem easy when you read about them or listen during class, but
practice is key to ensuring you understand the material. Practice
problems are provided for each lesson. The link above points you to the
appropriate link in the [course
notes](https://jsgosnell.github.io/cuny_biostats_book/){target="_blank"}.
You can make a copy (technically a *fork*, since you can't directly edit
it) of the entire course notes website in github \@
[https://github.com/jsgosnell/cuny_biostats_book](https://github.com/jsgosnell/cuny_biostats_book){target="_blank"}
and work from there. The benefit is this allows you to see updates to
the site (if you sync your fork). The downside is you have to work
interactively or build the entire site when you render a changed file.
This is doable but may take more time than students need (and may lead to merge issues!). 

Alternatively,your instructor may use a different
delivery method (like github classroom) or provide alternative problems.

**In general you should only work edit .qmd files!  Everything/anything else is produced during the session and should not be edited. All files can be uploaded to github though.** 

Solutions are also provided for all problems via the [course
notes](https://jsgosnell.github.io/cuny_biostats_book/){target="_blank"},
but try them before you look at the answers!

## Extra material

-   @davies1998

## Data referenced in class

-   @klem1989
