---
title: Introducing hypothesis testing via the binomial distribution
bibliography: ../references.bib
---

<!-- COMMENT NOT SHOW IN ANY OUTPUT: Code chunk below sets overall defaults for .qmd file; these inlcude showing output by default and looking for files relative to .Rpoj file, not .qmd file, which makes putting filesin different folders easier  -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Objective

To introduce students to hypothesis testing via the binomial test.By the
end of the lesson, you should be able to

-   define and explain null hypothesis significance testing

-   develop null and alternative hypotheses

-   define a p-value and explain how it may obtained and interpreted

-   define statistical power, type I error, and type II error

-   contrast one- and two-sided tests

-   define binomial data and evaluate using binomial tests

-   develop confidence intervals for binomial data

-   relate hypothesis testing and estimation methods

## Background reading

Course notes links for background reading also contain code used to
produce R output used in slides.

-   <https://jsgosnell.github.io/cuny_biostats_book/content/chapters/Binomial.html>

## [Lecture slides (click to open in Google slides!)](https://docs.google.com/presentation/d/1DNGtYoNSoG1ZKyy2E0uNAqQrPcczUweNyx4L9tIt3gU/edit?usp=sharing){target="_blank"}

```{=html}
<iframe src="https://docs.google.com/presentation/d/e/2PACX-1vQNeMNCd46WhyNcozIdsJqnA__HwMJfi6QXvGIa1koS3nsE444GLmQkwVpyvteT-JSsrxlQjYmuy53x/embed?start=false&loop=false&delayms=60000" frameborder="0" width="640" height="389" allowfullscreen="true" mozallowfullscreen="true" webkitallowfullscreen="true"></iframe>
```
## Connected swirl lesson

Swirl is an R package that provides guided lessons to help you learn and
review material. These lessons should serve as a bridge between all the
code provided in the slides and background reading and the key functions
and concepts from each lesson. A full course lesson (all lessons
combined) can also be downloaded using the following instructions.

**THIS IS ONE OF THE FEW TIMES I RECOMMEND WORKING DIRECTLY IN THE
CONSOLE! THERE IS NO NEED TO DEVELOP A SCRIPT FOR THESE INTERACTIVE
SESSIONS, THOUGH YOU CAN!**

-   install the "swirl" package

-   run the following code once on the computer to install a new course

    ```{r, eval=FALSE}
    library(swirl)
    install_course_github("jsgosnell", "JSG_swirl_lessons")
    ```

-   start swirl!

    ```{r, eval=F}
    swirl()
    ```

-   then follow the on-screen prompts to select the JSG_swirl_lessons
    course and the lessons you want

    -   Here we will focus on the **Hypothesis testing starting with
        binomial tests** lesson

-   TIP: If you are seeing duplicate courses (or odd versions of each),
    you can clear all courses and then re-download the courses by

    -   exiting swirl using escape key or bye() function

        ```{r, eval=F}
        bye()
        ```

    -   uninstalling and reinstalling courses

        ```{r, eval=F}
        uninstall_all_courses()
        install_course_github("jsgosnell", "JSG_swirl_lessons")
        ```

    -   when you restart swirl with swirl(), you may need to select

        -   No. Let me start something new

## [Connected assignment(click here)](https://jsgosnell.github.io/cuny_biostats_book/content/practice_problems/3_Introduction_to_hypothesis_testing_via_binomial_test.html){target="_blank"}

## Extra material

-   Tutorials produced by a group at UBC tha help visualize sampling,
    confidence intervals, and the central limit theorm
    -   [Sampling from a normally distributed
        population](https://www.zoology.ubc.ca/~whitlock/Kingfisher/SamplingNormal.htm){target="_blank"}
    -   [Confidence intervals for the
        mean](https://www.zoology.ubc.ca/~whitlock/Kingfisher/CIMean.htm){target="_blank"}
    -   [Sampling from a non-Normally distributed population (the
        central limit
        theorem)](https://www.zoology.ubc.ca/~whitlock/Kingfisher/CLT.htm){target="_blank"}
-   @perezgonzalez2015

## Data referenced in class

-   @klem1989

-   @bisazza1996

-   @geipel2020

-   @fischer2021

-   @hama2001
